Embodied Neuromorphic Benchmarks

### 2015-08-16 paper/paper.tex
Figures and finishing sections

Furthermore, using a physical robot imposes significant practical
difficulties when performing extensive benchmark testing.  When testing, we
often want to run the same task over and over again, both for robustness and
to see the effects of varying parameters.  With a physical robot, this means
manually setting up the task, letting the test run, gathering the resulting
data, and then resetting the robot back to the initial state.  Consequently,
issues like battery life become problematic, and not just because there is a
limited amount of time available for testing.  As the battery level changes,
the performance of the robot itself can also change.  Futhermore, for any
rigorous testing of the benchmark, we will want to examine situations where
the system fails.  This means that some of the testing will involve parameter
settings that lead to poor behaviour, which might have the undesireable 
result of causing physical damage.

@@

Given the success of this approach for evolutionary robotics, we propose
using minimal simulation for neuromorphic benchmarks.  We would argue that
one important use of a benchmark is \emph{generalization}.  That is, by knowing how well particular
hardware performs on a benchmark, you can reasonably infer how well
that hardware will perform in other situations.  For example, if an image recognition
algorithm performs well on the MNIST hand-written digit recognition
benchmark, this suggests that it may also perform well on
a different recognition task.  Of course, this inference will fail if that
algorithn has been specifically over-fit to that situtation.  For
that reason, it is useful to have benchmarks that covers a wide range
of variations on the task.  If the hardware performs well across that
variability, then it is more likely to also work in sufficiently similar new situations.

@@

It is important to note that there is a theoretical advantage to
using cheap robotics hardware for benchmarking, in addition to the practical
advantages.  In particular, we \emph{do not want benchmarks that rely on
high-speed, high-accuracy devices}.  The purpose of benchmarks
is not to indicate how well this neuromorphic hardware works to control this
particular robot in this task.  Rather, the purpose of a benchmark is to
characterise how well some specific neuromorphic hardware works on a task 
\emph{in general}.  The variability introduced in the minimal simulation means that the hardware
should be able to function across a wide variety of physical embodiments, and
so if we are to choose one particular physical embodiment to test in the
real world, then we should choose one that is not high-precision.
For this reason, we believe using cheap Lego robots is actually more informative
for benchmarking than expensive high-precision robots.\footnote{Of course,
for more complex benchmark tasks we may need sensory and motor capabiities
that are beyond that of a simple Lego robot.}

@@

However, this controller has difficulty in the presence of significant external
forces.  For example, consider a single motor controlling the angle of a single
arm.  If the arm is held out to the side, gravity acting on the mass of the
arm itself will pull the arm downward.  Thus to hold the arm still at a
particular $q_d$ will require the controller to apply a force to counteract
gravity.  Since the PD controller always produces an output $u=0$ when $q=q_d$,
it cannot compensate for gravity, and so the arm will stay stationary at some
angle below the desired angle. [TODO: add diagram]

The standard solution to this steady-state error is to add an I (integral) term ($K_i \int{(q_d-q) dt}$)
to the controller, making it a PID controller.  As the difference between where it is ($q$) and
where we want it to be ($q_d$) accumulates over time, the $K_i$ term will gradually increase the
the extra controlled force $u$ that is being applied until it is large enough to counteract
the external force of gravity (or whatever other external forces are present).
However, this approach has great difficulty when $q_d$ changes, since the
external force due to gravity changes depending on the position of the arm $q$.
The controller ends up having to ``relearn'' the
correct amount of extra force needed every time $q_d$ changes.

@@

Notice that we can think of this system as a three-layer neural
network, where the input and output layers are linear.  The first layer is $q$, the input state, one value for each joint.  The ``hidden'' layer
is the neurons producing activities $A$, the activity of a large number of neurons.  The output layer again has
one value per joint, and is the extra added signal to apply to the motors, $Ad$.
Given that this is a canonical example of the use of neural networks, we
expect that the majority of neuromorphic hardware is flexible enough to implement
this model.  Importantly, it functions well with spiking neuron models as 
well as non-spiking ones.  For spiking neurons, we consider $A$ to be the
instantaneous measure of the output of a neuron (i.e. whether or not it is currently outputing a spike),
filtered through a low-pass filter.  Further discussion of this sort of learning
rule and comparison to biological spiking neurons can be found in (\citealt{bekolay2013}).

@@

In addition, we add an external perturbing force.  In a real system,
this could be the effects of gravity given the current configuration of the
motors, or of other unexpected influences.  
Rather than chosing one particular fixed external force for our benchmark,
we \emph{randomly generate} this force each time the benchmark is run.  This
ensures that the benchmark covers a wide range of possible external forces
and motor configurations, rather than just one particular situation.

@@

For sensor noise, we note that the EV3 rotation encoders for the motors (the
devices that measure $q$) have a resolution of 0.0175 (1 degree).  This is a
very different sort of noise than the gaussian noise used in the simulation,
so we set the simulation noise $\sigma_q$ to be much larger (uniformly distributed
between 0 and 0.1).  Similarly, the motor resolution is 0.01, as it accepts 
integer values up to 100, so we set the motor noise $\sigma_u$ to be uniform between
0 and 0.1.

@@

For example, Figure \ref{fig:analysis_basic} shows the root-mean-squared
error (rmse) between $q$ and $q_d$ for three different neural systems.  For
this benchmark, $t_q$, $t_u$, $\tau_q$, and $\tau_u$ are chosen from $U(0, 0.01)$ (the uniform distribution), $\sigma_q$ and $\sigma_u$ are from $U(0, 0.1)$,
and $\beta$, $\gamma$, $\eta$, and $\zeta$ are all $N(0, 1)$.  As discussed
above, $K_f$ is 1 and $T$ is 10.  $q_d$ is set to be gaussian white noise with
a maximum frequency of 1Hz and RMS power of 1.  Each simulation is run
for 20 seconds, and the error is computed on the last 10 seconds.

The first hardware tested on this benchmark is an Intel i5-3337U CPU running
at 1.80GHz.  This is \emph{not} neuromorphic hardware, but serves as a
point of comparison.  The learning algorithm was implemented using Nengo,
a software toolkit for developing large-scale neural models that can be run
on various hardware platforms \cite{nengo}.  For the neuron model, we used
500 spiking Leaky-Integrate-and-Fire neurons.

@@

In Figure \ref{fig:analysis_basic}, all three systems perform equally well.
However, on that benchmark all three systems are implementing exactly 500
neurons.  The only thing that can be concluded from this is that the
differences in neuron model across that hardware does not significantly
impact performance.  A more interesting measure would be to compare based
on how many neurons each piece of hardware is capable of running in real time.
This is shown in Figure \ref{fig:analysis_compute}.  In this case, with the
current implementation, the CPU can
run 5200 neurons, the GPU 1500 neurons, and a single SpiNNaker core (of which
there are 18 on one chip) can run 500
neurons.  These values were measured empirically with the current versions
of the reference Nengo implementation, the Nengo OpenCL implementation,
and the Nengo SpiNNaker implementation (as of August 10, 2015).  
Figure \ref{fig:analysis_compute}
also shows the effects of varying the number of joints in the simulation,
giving an indication of the complexity of the external forces that can be
learned by this many neurons.

@@

While it is possible to run large neural models on standard CPUs and GPUs,
one of the primary advantages of neuromorphic hardware is its power
efficiency.  For this reason, the third benchmark normalizes the number of neurons
based on power consumption.  With a power budget of 1W per chip (with 16 used cores),
we estimate 0.0625W for the 500 neurons used here and round up to 0.1W to be
conservative.  Of course, neither the CPU nor the
GPU are designed to run on that little power.  For this reason,
on this benchmark we simply and scale it by the power consumption of the hardware.
For this power consumption we measure the difference between the idle power
consumption and the consumption when running the benchmark.  For the Intel
i5-3337U this was $34W - 11.5W = 12.5W$ and for the Nvidia Tesla C2075 GPU
this was $74W - 70W = 4W$.  This value is much lower than the peak power
consumption supported by the GPUi ($215W$), indicating that the current implementation
does not make extensive use of the GPU.  Indeed, initial analysis indicates that
the main bottleneck is communication between the GPU and the environment (i.e.,
the minimal simulation), and we feel it is appropriate that this benchmark
captures that limitation of the current GPU implementation.

Given the above considerations, we run the benchmark indicating that the CPU can run 42 neurons per 0.1W,
the GPU 38 neurons per 0.1W, and SpiNNaker can run 500 neurons per 0.1W.  If
this number of neurons is kept fixed and the number of joints controlled is
varied, we get the benchmark shown in Figure \ref{fig:analysis_power}.  This clearly
shows the power advantage of neuromorphic hardware over standard CPUs and
GPUs on this task.

@@

We can also use this sort of benchmark to examine the effects of various
parameters in the model.  For example, in Figure \ref{fig:analysis_delay_cpu}
we show what happens as the delays $t_q$ and $t_u$ are increased.
Importantly, the range on these parameters is larger than before ($U(0, 0.04)$
rather than $U(0, 0.01)$).  This shows that if the delay is short (less than 0.02), the controller performs well, and if it is very large (greater than 0.03),
the controller performs poorly.  However, for delays between 0.02 and 0.03,
the controller will sometimes perform well and sometimes perform poorly.  The
difference is due to the other random parameters in the system.  Interestingly,
running this same benchmark on SpiNNaker in Figure \ref{fig:analysis_delay_spinn} shows that SpiNNaker actually performs better on this
task.  This is somewhat surprising, as we are currently using the slow
Ethernet interface to SpiNNaker, rather than the high-speed I/O system that
is meant for motor control.  Further analysis is needed to determine exactly why this is.

@@

We also believe this approach to developing benchmarks would be applicable
to more complex and abstract tasks.  For example, much of the original work
on minimal simulation involved robot navigation and obstacle avoidance.  
\cite{Jakobi97evolutonaryrobotics} created a
minimal simulation of an environment with fixed walls, a mobile robot, and
various sensory systems.  A benchmark involving movement between locations
while avoiding collisions could be defined, and the environment could be
enhanced with moving obstacles and a richer set of sensors.  The eventual
goal would be to achieve reliable movement in an environment that includes
people.

@@

We thank Andrew Mundy for extensive work on the Nengo SpiNNaker backend, and
James Knight for his SpiNNaker implementation of the learning rule used here.

@@

@@

%\begin{figure}
%\begin{center}
%\includegraphics[width=10cm]{logo2}% This is an *.eps file
%\end{center}
%\textbf{\refstepcounter{figure}\label{fig:02} Figure \arabic{figure}.}{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures }
%\end{figure}

>>>>

Furthermore, using a physical robot imposes significant practical
difficulties when performing extensive benchmark testing.  When testing, we
often want to run the same task over and over again, both for robustness and
to see the effects of varying parameters.  With a physical robot, this means
manually setting up the task, letting the test run, gathering the resulting
data, and then resetting the robot back to the initial state.  Consequently,
issues like battery life become problematic, and not just because there is a
limited amount of time available for testing.  As the battery level changes,
the performance of the robot itself can also change.  Futhermore, for any
rigorous testing of the benchmark, we will want to examine situations where
the system fails.  This means that some of the testing will involve parameter
settings that lead to poor behaviour, which might have the undesirable 
result of causing physical damage.

@@

Given the success of this approach for evolutionary robotics, we propose
using minimal simulation for neuromorphic benchmarks.  We would argue that
one important use of a benchmark is \emph{generalization}.  That is, by knowing how well particular
hardware performs on a benchmark, you can reasonably infer how well
that hardware will perform in other situations.  For example, if an image recognition
algorithm performs well on the MNIST hand-written digit recognition
benchmark, this suggests that it may also perform well on
a different recognition task.  Of course, this inference will fail if that
algorithm has been specifically over-fit to that situation.  For
that reason, it is useful to have benchmarks that covers a wide range
of variations on the task.  If the hardware performs well across that
variability, then it is more likely to also work in sufficiently similar new situations.

@@

It is important to note that there is a theoretical advantage to
using cheap robotics hardware for benchmarking, in addition to the practical
advantages.  In particular, we \emph{do not want benchmarks that rely on
high-speed, high-accuracy devices}.  The purpose of benchmarks
is not to indicate how well this neuromorphic hardware works to control this
particular robot in this task.  Rather, the purpose of a benchmark is to
characterise how well some specific neuromorphic hardware works on a task 
\emph{in general}.  The variability introduced in the minimal simulation means that the hardware
should be able to function across a wide variety of physical embodiments, and
so if we are to choose one particular physical embodiment to test in the
real world, then we should choose one that is not high-precision.
For this reason, we believe using cheap Lego robots is actually more informative
for benchmarking than expensive high-precision robots.\footnote{Of course,
for more complex benchmark tasks we may need sensory and motor capabilities
that are beyond that of a simple Lego robot.}

@@

However, this controller has difficulty in the presence of significant external
forces.  For example, consider a single motor controlling the angle of a single
arm.  If the arm is held out to the side, gravity acting on the mass of the
arm itself will pull the arm downward.  Thus to hold the arm still at a
particular $q_d$ will require the controller to apply a force to counteract
gravity.  Since the PD controller always produces an output $u=0$ when $q=q_d$,
it cannot compensate for gravity, and so the arm will stay stationary at some
angle below the desired angle (Figure \ref{fig:system}).

The standard solution to this steady-state error is to add an I (integral) term ($K_i \int{(q_d-q) dt}$)
to the controller, making it a PID controller.  As the difference between where it is ($q$) and
where we want it to be ($q_d$) accumulates over time, the $K_i$ term will gradually increase
the extra controlled force $u$ that is being applied until it is large enough to counteract
the external force of gravity (or whatever other external forces are present).
However, this approach has great difficulty when $q_d$ changes, since the
external force due to gravity changes depending on the position of the arm $q$.
The controller ends up having to ``relearn'' the
correct amount of extra force needed every time $q_d$ changes.

@@

Notice that we can think of this system as a three-layer neural
network, where the input and output layers are linear.  The first layer is $q$, the input state, one value for each joint.  The ``hidden'' layer
is the neurons producing activities $A$, the activity of a large number of neurons.  The output layer again has
one value per joint, and is the extra added signal to apply to the motors, $Ad$.
Given that this is a canonical example of the use of neural networks, we
expect that the majority of neuromorphic hardware is flexible enough to implement
this model.  Importantly, it functions well with spiking neuron models as 
well as non-spiking ones.  For spiking neurons, we consider $A$ to be the
instantaneous measure of the output of a neuron (i.e. whether or not it is currently outputting a spike),
filtered through a low-pass filter.  Further discussion of this sort of learning
rule and comparison to biological spiking neurons can be found in (\citealt{bekolay2013}).

@@

In addition, we add an external perturbing force.  In a real system,
this could be the effects of gravity given the current configuration of the
motors, or of other unexpected influences.  
Rather than choosing one particular fixed external force for our benchmark,
we \emph{randomly generate} this force each time the benchmark is run.  This
ensures that the benchmark covers a wide range of possible external forces
and motor configurations, rather than just one particular situation.

@@

For sensor noise, we note that the EV3 rotation encoders for the motors (the
devices that measure $q$) have a resolution of 0.0175 (1 degree).  This is a
very different sort of noise than the Gaussian noise used in the simulation,
so we set the simulation noise $\sigma_q$ to be much larger (uniformly distributed
between 0 and 0.1).  Similarly, the motor resolution is 0.01, as it accepts 
integer values up to 100, so we set the motor noise $\sigma_u$ to be uniform between
0 and 0.1.

@@

For example, Figure \ref{fig:analysis_basic} shows the root-mean-squared
error (rmse) between $q$ and $q_d$ for three different neural systems.  For
this benchmark, $t_q$, $t_u$, $\tau_q$, and $\tau_u$ are chosen from $U(0, 0.01)$ (the uniform distribution), $\sigma_q$ and $\sigma_u$ are from $U(0, 0.1)$,
and $\beta$, $\gamma$, $\eta$, and $\zeta$ are all $N(0, 1)$.  As discussed
above, $K_f$ is 1 and $T$ is 10.  $q_d$ is set to be Gaussian white noise with
a maximum frequency of 1Hz and RMS power of 1.  Each simulation is run
for 20 seconds, and the error is computed on the last 10 seconds.

The first hardware tested on this benchmark is an Intel i5-3337U CPU running
at 1.80GHz.  This is not neuromorphic hardware, but serves as a
point of comparison.  The learning algorithm was implemented using Nengo,
a software toolkit for developing large-scale neural models that can be run
on various hardware platforms \cite{nengo}.  For the neuron model, we used
500 spiking Leaky-Integrate-and-Fire neurons.

@@

In Figure \ref{fig:analysis_basic}, all three systems perform equally well.
However, on that benchmark all three systems are implementing exactly 500
neurons.  The only thing that can be concluded from this is that the
differences in neuron model across that hardware does not significantly
impact performance.  A more interesting measure would be to compare based
on how many neurons each piece of hardware is capable of running in real time.
This is shown in Figure \ref{fig:analysis_compute}.  In this case, with the
current implementation, the CPU can
run 5200 neurons, the GPU 1500 neurons, and a single SpiNNaker core (of which
there are 18 on one chip) can run 500
neurons.  These values were measured empirically with the current versions
of the reference Nengo implementation, the Nengo OpenCL implementation,
and the Nengo SpiNNaker implementation (as of August 10, 2015).  All other
parameters are as before.

@@

While it is possible to run large neural models on standard CPUs and GPUs,
one of the primary advantages of neuromorphic hardware is its power
efficiency.  For this reason, the third benchmark normalizes the number of neurons
based on power consumption.  With a power budget of 1W per chip (with 16 used cores),
we estimate 0.0625W for the 500 neurons used here and round up to 0.1W to be
conservative.  Of course, neither the CPU nor the
GPU are designed to run on that little power.  For this reason,
on this benchmark we simply and scale it by the power consumption of the hardware.
For this power consumption we measure the difference between the idle power
consumption and the consumption when running the benchmark.  For the Intel
i5-3337U this was $34W - 11.5W = 22.5W$ and for the Nvidia Tesla C2075 GPU
this was $74W - 70W = 4W$.  This value is much lower than the peak power
consumption supported by the GPU ($215W$), indicating that the current implementation
does not make extensive use of the GPU.  Indeed, initial analysis indicates that
the main bottleneck is communication between the GPU and the environment (i.e.,
the minimal simulation), and we feel it is appropriate that this benchmark
captures that limitation of the current GPU implementation.

Given the above considerations, we run the benchmark indicating that the CPU can run 23 neurons per 0.1W,
the GPU 38 neurons per 0.1W, and SpiNNaker can run 500 neurons per 0.1W.  
As shown in Figure \ref{fig:analysis_power}, the neuromorphic hardware
outperforms the CPU and GPU on this task.

@@

We can also use this sort of benchmark to examine the effects of various
parameters in the model.  For example, Figure \ref{fig:analysis_delay}
shows what happens as the delays $t_q$ and $t_u$ are increased.
Importantly, the range on these parameters is larger than before ($U(0, 0.04)$
rather than $U(0, 0.01)$).  This shows that if the delay is short (less than 0.02), the controller performs well, and if it is very large (greater than 0.03),
the controller performs poorly.  However, for delays between 0.02 and 0.03,
the controller will sometimes perform well and sometimes perform poorly.  The
difference is due to the other random parameters in the system.  Interestingly,
running SpiNNaker actually performs better on this
task than the CPU (the GPU data is equivalent to the CPU and is not shown).  
This is somewhat surprising, as we are currently using the slow
Ethernet interface to SpiNNaker, rather than the high-speed I/O system that
is meant for motor control.  Further analysis is needed to determine exactly why this is.

@@

We also believe this approach to developing benchmarks would be applicable
to more complex and abstract tasks.  For example, much of the original work
on minimal simulation involved robot navigation and obstacle avoidance.  
\cite{Jakobi97evolutionaryrobotics} created a
minimal simulation of an environment with fixed walls, a mobile robot, and
various sensory systems.  A benchmark involving movement between locations
while avoiding collisions could be defined, and the environment could be
enhanced with moving obstacles and a richer set of sensors.  The eventual
goal would be to achieve reliable movement in an environment that includes
people.

@@

We thank Andrew Mundy for extensive work on the Nengo SpiNNaker backend, and
James Knight for his prototype SpiNNaker implementation of the learning rule used here.

@@

\begin{figure}[h!]
\begin{center}
\includegraphics[width=10cm]{figures/SingleArmSystem}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:system} Figure \arabic{figure}.}{ 
     An abstract physical system that could be controlled.  In order to hold
     the joint $q$ at a desired angle $q_d$, a force must be applied to
     counteract the pull of gravity.  The magnitude of this force is a
 function of $q$ which can be learned.}
\end{figure}

@@

\begin{figure}[h!]
\begin{center}
\includegraphics[width=18cm]{figures/plot_basic}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:analysis_basic} Figure \arabic{figure}.}{ Benchmark results comparing three hardware systems.
     Each system is running 500 neurons.  The hardware do not statistically
     significantly differ, and are all clear improvements over the having
     no adaptation at all.
     Scatterplots show individual runs, the 
     shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=18cm]{figures/plot_compute}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:analysis_compute} Figure \arabic{figure}.}{ Benchmark results comparing three hardware systems
     in terms of their performance when running as many neurons as they are capable of in real time.  Scatterplots show individual runs, the 
     shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.  In this case,
     all three hardware systems perform equally well.}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=18cm]{figures/plot_power}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:analysis_power} Figure \arabic{figure}.}{ Benchmark results comparing three hardware systems
     in terms of their performance when running as many neurons as they are capable of per 0.1 Watt of power consumption.  Scatterplots show individual runs, the 
     shaded area is the mean plus or minus the standard deviation, and the 95\% confidence interval of the mean is shown.  The SpiNNaker system is a statistically
 significant improvement over the other two.}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=18cm]{figures/plot_delay}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:analysis_delay} Figure \arabic{figure}.}{ Benchmark results comparing the effects
     of communication delay for the CPU and SpiNNaker systems.  The delays $t_q$ and $t_u$ are randomly varied.  Shaded
 area is the mean plus or minus one standard deviation, smoothed with a Gaussian kernel of $\sigma=0.005$.}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=18cm]{figures/plot_vary_neurons}
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:analysis_neurons_cpu} Figure \arabic{figure}.}{ Benchmark results examining the relationship
     between number of neurons and the number of simulated joints $N$.  The
     benchmark was run on the Intel i5-3337U CPU.  Shaded
 area is the mean plus or minus one standard deviation, smoothed with a Gaussian kernel of $\sigma=1$ in the $log_2$ domain.}
\end{figure}
%\begin{figure}
%\begin{center}
%\includegraphics[width=10cm]{logo2}% This is an *.eps file
%\end{center}
%\textbf{\refstepcounter{figure}\label{fig:02} Figure \arabic{figure}.}{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures }
%\end{figure}
